DRAFT Example Kitsap County Pedestrian Facilities Metrics/Evaluation Report

### 2022-10-06 Draft Ped Metrics Example.Rmd
clean up tables

@@

These are the roads that will be evaluated.
```{r}
library(tidyverse)
library(sf)
# groupcl <- ncdf %>% #clean_roadcl %>%  
#   st_set_geometry(., NULL) %>% 
#   group_by(RUCODE, road_class) %>% 
#   summarise(roads = nice_comma(n_distinct(FULL_NAME)), 
#             segments = nice_comma(n()), 
#             length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280,1),
#             .groups = "drop") %>%
#   mutate(RUCODE = factor(RUCODE, 
#                          levels = c(0,1,2), 
#                          labels = c("Unknown", "Rural", "Small Urban")))
# 
# colnames(groupcl) <- c("Rural/Urban", "Road Classification", 
#                        "Count of Roads", "Count of Segments", 
#                        "Length in Miles") 
groupcl <- ncdf %>% #clean_roadcl %>%  
  st_set_geometry(., NULL) %>% 
  group_by(road_class) %>% 
  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
            segments = nice_comma(n(),0), 
            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
            .groups = "drop") 
df <- ncdf %>% #clean_roadcl %>%  
  st_set_geometry(., NULL) %>% 
  group_by(road_class = "Total") %>% 
  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
            segments = nice_comma(n(),0), 
            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
            .groups = "drop") 
groupcl <- bind_rows(groupcl, df)

colnames(groupcl) <- c("Road Classification", 
                       "Count of Roads", "Count of Segments", 
                       "Length in Miles") 
knitr::kable(groupcl, 
             caption = "Roads by Class in Kitsap County, excluding cities, highways, and easements")
```

@@

knitr::kable(df, 
             caption = "Miles of Sidewalks by Width of Sidewalk (single side)")
```

@@

knitr::kable(df, 
             caption = "Miles of Paved Shoulder by Width of Shoulder (single side)")
```

@@

```{r combine-rd-sw-sh, echo = TRUE}
# combine the sidewalk and shoulder and roads data
x <- st_as_sf(nc_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
x$RD_LOG_ID <- 43809
comb <- st_intersection(select(ncdf, RD_LOG_ID_R = RD_LOG_ID, 
                                     SEGMENT_ID_R = SEGMENT_ID,
                                     FULL_NAME_R = FULL_NAME, 
                                     LENGTH_R = LENGTH,
                                     road_class, RUCODE), 
                             st_buffer(bind_rows(nc_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
  filter(RD_LOG_ID == RD_LOG_ID_R) %>%
  distinct(RD_LOG_ID_R, SEGMENT_ID_R, RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, .keep_all = TRUE) %>%
  mutate(distance = as.integer(st_length(.)))

#distance is the length of the matched road segment and sidewalk or shoulder 
#comb %>% st_set_geometry(., NULL) %>% 
#  group_by(road_class, SIDE_OF_RO, TYPE) %>%
#  summarize(s_miles = sum(distance)/5280,
#            miles = format(round(as.numeric(sum(distance)/5280), 1), 
#                           digits = 1, nsmall=1, big.mark=","),
#            .groups = "drop")
```

@@

```{r tbl_standards}
# Road Standards Tables 3.3 and 3.4
library(tidyverse)
library(kableExtra)
road_standards_given <- tribble(
  ~setting, ~class, ~characteristics, ~sidewalk, ~shoulder, ~note,
  "Urban", "Local Road", NA, 5, NA,NA,
  "Urban", "Local Sub-collector", NA, 5, NA,"sub-collector is not defined in road data",
  "Urban", "Arterial", NA, 6, NA,NA,
  "Urban", "Collector", NA, 6, NA,NA,
  "Rural", "Local Road", NA, NA, 3,NA,
  "Rural", "Local Sub-collector", NA, NA, 4,"sub-collector is not defined in road data",
  "Rural", "Collector", "ADT 400-750", NA, 3,"ADT not defined in road data",
  "Rural", "Collector", "ADT 751-1000", NA, 4,"ADT not defined in road data",
  "Rural", "Collector", "DHV 100-200", NA, 6,"DHV not defined in road data",
  "Rural", "Collector", "DHV > 200", NA, 8,"DHV not defined in road data",
  "Rural", "Minor Arterial", "DHV < 100", NA, 4,"DHV not defined in road data",
  "Rural", "Minor Arterial", "DHV 100-200", NA, 6,"DHV not defined in road data",
  "Rural", "Minor Arterial", "DHV > 200", NA, 8,"DHV not defined in road data",
  "Rural", "Principal Arterial", "DHV <200", NA, 6,"DHV not defined in road data",
  "Rural", "Principal Arterial", "DHV > 200", NA, 8,"DHV not defined in road data",
)

knitr::kable(road_standards_given,
             caption = "Summary of Tables 3.3 and 3.4 in Kitsap County Road Standards [@kc_road_standards]")  %>%
   footnote(general = "Within the road standards, the minimum shoulder width may be reduced to the minimum required by AASHTO, which would be 4-8 feet, depending on the ADT")
```

@@

knitr::kable(road_standards,
             caption = "Simplified Assumptions to Allow for Evaluation")  %>%
   footnote(general = "This is only a simplification to allow for an evaluation without knowledge of ADT or DHV on the roads under evaluation")
```

@@

|Segment|Segment Score | Segment Centerline Linear Feet |
|------|:-----|:-----|
|A|6|100|
|B|0|10|
|C|10|50|
|D|8|200|
|*Total*| *?* | *360*|

@@

knitr::kable(score_interpretation,
             caption = "Interpretation of Scores for Segments and for Areas")  %>%
   footnote(general = "For a segment, adequate means one side is good or both sides meet AASHTO minimums and minimal means one side meets AASHTO minimums")
```

@@

```{r}
knitr::kable(comb_score_by_setting_class)
```

More detailed information
```{r}
knitr::kable(
  comb_score2 %>%
  group_by(setting, road_class, score) %>% 
  summarise(roads = nice_comma(n_distinct(RD_LOG_ID_R)), 
            segments = nice_comma(n()), 
            #road_len = nice_comma(sum(the_len, na.rm=TRUE)/5280, 1),
            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
            .groups = "drop") 
  )
```

@@

knitr::kable(head(roads_0))

@@

```{r}
roads_10 <- comb_score2 %>%
  filter(score == 10) %>%
  group_by(setting, road_class, "Road Name" = FULL_NAME_R) %>%
  summarise(segments = nice_comma(n()), 
            road_len = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
            .groups = "drop")

knitr::kable(roads_10)
```

>>>>

library(kableExtra)

@@

These are the roads that will be evaluated.
```{r}
library(tidyverse)
library(sf)

groupcl <- ncdf %>%  
  st_set_geometry(., NULL) %>% 
  group_by(road_class) %>% 
  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
            segments = nice_comma(n(),0), 
            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
            .groups = "drop") 
df <- ncdf %>% #clean_roadcl %>%  
  st_set_geometry(., NULL) %>% 
  group_by(road_class = "Total") %>% 
  summarise(roads = nice_comma(n_distinct(FULL_NAME),0), 
            segments = nice_comma(n(),0), 
            length = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
            .groups = "drop") 
groupcl <- bind_rows(groupcl, df)

colnames(groupcl) <- c("Road Classification", 
                       "Count of Roads", "Count of Segments", 
                       "Length in Miles") 
knitr::kable(groupcl, 
             caption = "Roads by Class in Kitsap County, excluding cities, highways, and easements")  %>%
  kable_material(c("striped", "hover"))
```

@@

knitr::kable(df, 
             caption = "Miles of Sidewalks by Width of Sidewalk (single side)") %>%
  kable_material(c("striped", "hover"))
```

@@

knitr::kable(df, 
             caption = "Miles of Paved Shoulder by Width of Shoulder (single side)") %>%
  kable_material(c("striped", "hover"))
```

@@

```{r combine-rd-sw-sh}
# combine the sidewalk and shoulder and roads data
x <- st_as_sf(nc_sw_sh) %>% mutate(LENGTH = as.integer(LENGTH)) %>% filter(RD_LOG_ID == 49435 & LENGTH > 1000)
x$RD_LOG_ID <- 43809
comb <- st_intersection(select(ncdf, RD_LOG_ID_R = RD_LOG_ID, 
                                     SEGMENT_ID_R = SEGMENT_ID,
                                     FULL_NAME_R = FULL_NAME, 
                                     LENGTH_R = LENGTH,
                                     road_class, RUCODE), 
                             st_buffer(bind_rows(nc_sw_sh, x), 50, endCapStyle="FLAT")) %>% 
  filter(RD_LOG_ID == RD_LOG_ID_R) %>%
  distinct(RD_LOG_ID_R, SEGMENT_ID_R, RD_LOG_ID, SEGMENT_ID, SIDE_OF_RO, .keep_all = TRUE) %>%
  mutate(distance = as.integer(st_length(.)))

@@

```{r tbl_standards}
# Road Standards Tables 3.3 and 3.4
library(tidyverse)
library(kableExtra)
road_standards_given <- tribble(
  ~setting, ~class, ~characteristics, ~sidewalk, ~shoulder, 
  "Urban", "Local Road", NA, 5, NA,
  "Urban", "Local Sub-collector", NA, 5, NA,
  "Urban", "Arterial", NA, 6, NA,
  "Urban", "Collector", NA, 6, NA,
  "Rural", "Local Road", NA, NA, 3,
  "Rural", "Local Sub-collector", NA, NA, 4,
  "Rural", "Collector", "ADT 400-750", NA, 3,
  "Rural", "Collector", "ADT 751-1000", NA, 4,
  "Rural", "Collector", "DHV 100-200", NA, 6,
  "Rural", "Collector", "DHV > 200", NA, 8,
  "Rural", "Minor Arterial", "DHV < 100", NA, 4,
  "Rural", "Minor Arterial", "DHV 100-200", NA, 6,
  "Rural", "Minor Arterial", "DHV > 200", NA, 8,
  "Rural", "Principal Arterial", "DHV <200", NA, 6,
  "Rural", "Principal Arterial", "DHV > 200", NA, 8,
)

knitr::kable(road_standards_given,
             caption = "Summary of Tables 3.3 and 3.4 in Kitsap County Road Standards [@kc_road_standards]")  %>%
   footnote(general = "Within the road standards, the minimum shoulder width may be reduced to the minimum required by AASHTO, which would be 4-8 feet, depending on the ADT. Note: sub-collector, ADT, and DHV are not defined in the publicly available road data.") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

@@

knitr::kable(road_standards,
             caption = "Simplified Assumptions to Allow for Evaluation")  %>%
   footnote(general = "This is only a simplification to allow for an evaluation without knowledge of ADT or DHV on the roads under evaluation") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

@@

```{r}
example_tbl <- tribble(
  ~Segment, ~"Segment Score", ~"Segment Centerline Linear Feet",
  "A", 6, 100,
  "B", 0, 10,
  "C",10,50,
  "D",8,200,
  "Total",7.5 ,360,
)

knitr::kable(example_tbl) %>%
kable_styling(bootstrap_options = "striped", full_width = F, position = "float_left")
```

@@

knitr::kable(score_interpretation,
             caption = "Interpretation of Scores for Segments and for Areas")  %>%
   footnote(general = "For a segment, adequate means one side is good or both sides meet AASHTO minimums and minimal means one side meets AASHTO minimums") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

@@

Looking by the same breakdown that we have for standards, we can see where the scores are highest and lowest.

```{r}
knitr::kable(comb_score_by_setting_class) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

```{r}
# More detailed information
# knitr::kable(
#   comb_score2 %>%
#   group_by(setting, road_class, score) %>% 
#   summarise(roads = nice_comma(n_distinct(RD_LOG_ID_R)), 
#             segments = nice_comma(n()), 
#             #road_len = nice_comma(sum(the_len, na.rm=TRUE)/5280, 1),
#             dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
#             .groups = "drop") 
#   )
```

@@

knitr::kable(head(roads_0)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

@@

Several road segments were able to achieve a score of 10 - meaning that the segment met the standard on both sides of the road. A few details for roads where at least one segment has a 10 are shown here. 
```{r}
roads_10 <- comb_score2 %>%
  filter(score == 10) %>%
  group_by(setting, road_class, "Road Name" = FULL_NAME_R) %>%
  summarise(segments = nice_comma(n()), 
            road_len = nice_comma(sum(LENGTH, na.rm=TRUE)/5280, 1),
            dist = nice_comma(sum(distance, na.rm=TRUE)/5280, 1),
            .groups = "drop")

knitr::kable(roads_10) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```
